---
title: "Fragility reproducibility results for patient 01 and 26 seizure 1"
author: "Anne-Cecile Lesage, PhD, Oliver Zhou, BA, Jiefei Wang, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Fragility reproducibility results for patient 01 and 26 seizure 1}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
library(EZFragility)
knitr::opts_chunk$set(
    collapse = TRUE,
    comment = "#>",
    fig.align = "center",
    fig.width = 10,
    fig.height = 8
)
```

This vignette shows how to reproduce results from the companion software paper of the <a href="https://cran.r-project.org/web/packages/fragility/index.html" style="color:blue; text-decoration:none;"> EZFragility CRAN R package</a> for the first seizure of patients 01 and 26. This allow to check how well the EZFragility package reproduces results from the <a href="https://www.nature.com/articles/s41593-021-00901-w" style="color:blue; text-decoration:none;"> original neural fragility paper </a>(see Figure 4 and extended seizure 4)


## Step 1.  Visualize content of data package

To allow fast and easy exploration of fragility results,  we preprocessed with <a href="https://rave.wiki/" style="color:blue; text-decoration:none;"> RAVE2 </a> the multipatient data from 
the openneuro public data set <a href="https://openneuro.org/datasets/ds003029/versions/1.0.0/metadata" style="color:blue; text-decoration:none;"> Openneuro Fragility Data Set</a>. Main steps include notch filtering to remove power line interference, re-referencing, epoching based on seizure onset time to isolate relevant segments of [-30:30s]. The data names list (dl) is displaying the patient identifier and the seizure number of each Epoch object file ( <a href="https://github.com/Jiefei-Wang/Epoch" style="color:blue; text-decoration:none;"> Epoch package</a>).
```{r}
dl<-EpochDownloader()
names(dl)
```


## -----------------------
## Patient 01
## -----------------------

## Step 2.  Download Epoch object for patient 01 seizure 1
The preprocessed voltage data from patient pt01 seizure 1 can be loaded in the window [-30:30]s as an Epoch package object and cropped in the [-10:10]s time window around seizure onset with the following code.
```{r}
pt01sz1 <- dl$Retrostudy_subpt01_1
pt01sz1Clipped <- crop(pt01sz1, start=-10, end=10)
pt01sz1Clipped 
```
Use tblData, rowData, colData, metaData to get the data as directed in <a href="https://github.com/Jiefei-Wang/Epoch" style="color:blue; text-decoration:none;"> Epoch package</a>.

## Step 3.  Visualize voltage plot
sozIndex is the index of the electrodes which are labeled as in the SOZ in the original paper. The following code reproduce the voltage plot for a display subset of electrodes. The SOZ electrode names are highlighted in red as in the original Figure 4.
```{r}
sozIndex <- which(rowData(pt01sz1Clipped)$soz==TRUE)
display <- c(sozIndex, 77:80)
sz_onset<-0
vplotpt01sz1<-visuIEEGData(epoch = pt01sz1Clipped[display, ])

elecColor <- rep("black", length(display))
elecColor[1:length(sozIndex)] <- 'red'

elecColor <- rev(elecColor) # match the electrode order in the plot

vplotpt01sz1<- vplotpt01sz1 + ggplot2::geom_vline(xintercept = sz_onset, color = "black", linetype = "dashed", linewidth = 1)

vplotpt01sz1<- vplotpt01sz1 + ggplot2::theme(
  
  axis.text.y = ggtext::element_markdown(colour = elecColor)
  
)
vplotpt01sz1
```

## Step 4.  Compute and plot the fragility heatmap

The following code computes the fragility matrix on all electrodes and store the results in the Fragility class object pt01sz1Frag
```{r}
library(doSNOW)
# compute fragility

cl <- makeCluster(4, type = "SOCK")
registerDoSNOW(cl)

windowNum <- 250
step <- 125
pt01sz1Frag <- calcAdjFrag(epoch = pt01sz1Clipped, window = windowNum, step = step, parallel = TRUE, nSearch=100L,progress = TRUE)

# Stop the parallel backend
stopCluster(cl)
```

The following code plots the fragility heatmap with the same display options as the previous voltage plot. Looking at both plots allows to check correlation between soz patterns
on electrode voltage and high fragility. PT01 was a surgical success, with the high-fragility electrodes identified by the algorithm matching those identified by clinical epileptologists.
```{r}
elecNames  <- pt01sz1Frag$electrodes
startTimes <- pt01sz1Frag$startTimes

indexsz<-which(abs(startTimes)<=0.01)



#display <- c(pt01sozNames, "MLT1", "MLT2", "MLT3", "MLT4")
fplotpt01sz1<-plotFragHeatmap(frag = pt01sz1Frag[display], groupIndex = c(1:10),ranked=FALSE)
fplotpt01sz1<-fplotpt01sz1 + ggplot2::geom_vline(xintercept = indexsz, color = "black", linetype = "dashed", linewidth = 1)
fplotpt01sz1<-fplotpt01sz1 + ggplot2::theme(
  
  axis.text.y = ggtext::element_markdown(colour = elecColor)
  
)
fplotpt01sz1
```

Setting the "ranked" variable to TRUE in the plotFragHeatmap allows more contrast and render the heatmap closer to original figure 4.

```{r}
fplotpt01sz1<-plotFragHeatmap(frag = pt01sz1Frag[display], groupIndex = c(1:10),ranked=TRUE)
fplotpt01sz1<-fplotpt01sz1 + ggplot2::geom_vline(xintercept = indexsz, color = "black", linetype = "dashed", linewidth = 1)
fplotpt01sz1<-fplotpt01sz1 + ggplot2::theme(
  
  axis.text.y = ggtext::element_markdown(colour = elecColor)
  
)
fplotpt01sz1
```

## Step 5.  Compute mean and standard deviation statistics for SOZ and non SOZ electrode group


The following code allows to reproduce the mean and standard visualization statistics of two electrode groups (soz labeled electrodes and non soz labeled electrodes)
as in extended figure 4 from the original paper. This plot shows that the fragility biomarker statistics are significantly higher in the soz labeled group correlated with the ground truth success outcome for this patient. 

```{r}
fdistribpt01sz1<-plotFragDistribution(frag = pt01sz1Frag, groupIndex = sozIndex, bandType="SD", rollingWindow = 1,ranked=FALSE) ## p 1481 Extended Fig.4
fdistribpt01sz1<-fdistribpt01sz1 + ggplot2::geom_vline(xintercept = sz_onset, color = "black", linetype = "dashed", linewidth = 1)
fdistribpt01sz1
```

## -----------------------
## Patient 26
## -----------------------

## Step 2.  Download Epoch object for patient 01 seizure 1
The preprocessed voltage data from patient pt26 seizure 1 can be loaded in the window [-30:30]s as an Epoch package object and cropped in the [-10:10]s time window around seizure onset with the following code.
```{r}
pt26sz1 <- dl$Retrostudy_subjh103_1
pt26sz1Clipped <- crop(pt26sz1, start=-10, end=10)
pt26sz1Clipped 
```

## Step 3.  Visualize voltage plot 
The same visualization process is repeated here for patient 26 as in the original Figure 4.
```{r}
sozIndex <- which(rowData(pt26sz1)$soz==TRUE)
sozIndex<-sozIndex[-c(17,18,19,20)]

display <- c(c(1:2),sozIndex[1:16], c(53:56),sozIndex[17:18])

sz_onset<-0
vplotpt26sz1<-visuIEEGData(epoch = pt26sz1Clipped[display, ])

elecColor <- rep("black", length(display))
elecColor[1:length(sozIndex)] <- 'red'

elecColor <- rev(elecColor) # match the electrode order in the plot

vplotpt26sz1<- vplotpt26sz1 + ggplot2::geom_vline(xintercept = sz_onset, color = "black", linetype = "dashed", linewidth = 1)

vplotpt26sz1<- vplotpt26sz1 + ggplot2::theme(
  
  axis.text.y = ggtext::element_markdown(colour = elecColor)
  
)
vplotpt26sz1
```

## Step 4.  Compute and plot the fragility heatmap
The following code computes the fragility matrix on all electrodes and store the results in the Fragility class object pt26sz1Frag
```{r}
library(doSNOW)
# compute fragility

cl <- makeCluster(4, type = "SOCK")
registerDoSNOW(cl)

windowNum <- 250
step <- 125
pt26sz1Frag <- calcAdjFrag(epoch = pt26sz1Clipped, window = windowNum, step = step, parallel = TRUE, nSearch=100L,progress = TRUE)

# Stop the parallel backend
stopCluster(cl)
```

The following code plots the fragility heatmap with the same display options as the previous voltage plot. PT26 was a surgical failure, and the high-fragility electrodes were not consistent with clinical EEG interpretation
```{r}
elecNames  <- pt26sz1Frag$electrodes
startTimes <- pt26sz1Frag$startTimes

indexsz<-which(abs(startTimes)<=0.01)



#display <- c(pt26sozNames, "MLT1", "MLT2", "MLT3", "MLT4")
fplotpt26sz1<-plotFragHeatmap(frag = pt26sz1Frag[display], groupIndex = c(1:10),ranked=FALSE)
fplotpt26sz1<-fplotpt26sz1 + ggplot2::geom_vline(xintercept = indexsz, color = "black", linetype = "dashed", linewidth = 1)
fplotpt26sz1<-fplotpt26sz1 + ggplot2::theme(
  
  axis.text.y = ggtext::element_markdown(colour = elecColor)
  
)
fplotpt26sz1
```

## Step 5.  Compute mean and standard deviation statistics for SOZ and non SOZ electrode group 

The following code allows to reproduce the mean and standard visualization statistics of two electrode groups (soz labeled electrodes and non soz labeled electrodes)
as in extended figure 4 from the original paper. This plot shows that the fragility biomarker statistics are not significantly higher in the soz labeled group correlated with the ground truth failure outcome for this patient. 

```{r}
fdistribpt26sz1<-plotFragDistribution(frag = pt26sz1Frag, groupIndex = sozIndex, bandType="SD", rollingWindow = 1,ranked=FALSE) ## p 1481 Extended Fig.4
fdistribpt26sz1<-fdistribpt26sz1 + ggplot2::geom_vline(xintercept = sz_onset, color = "black", linetype = "dashed", linewidth = 1)
fdistribpt26sz1
```